---
title: "BenchmarkingFinal"
author: "Logan Wright"
date: "July 15, 2019"
output: pdf_document
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
options(stringsAsFactors = F)
options(scipen=999)
options(digits=8)
library("plyr")
library("genefilter")
library(ROTS); library(data.table); library(pwr); library(EnhancedVolcano)
library(limma); library(PECA); library(readxl); library(ztable)
library(dplyr); library(tidyr); library(ggplot2); library(ggthemes); 
library(printr); library(mratios); library(extrafont); library(qvalue); library(cowplot); library(gridExtra)
knitr::opts_chunk$set(echo = TRUE)

#BiocManager::install(c("ROTS", "pwr" , "limma" , "EnhancedVolcano" , "PECA" , "readxl" , "ztable" , "mratios" , "qvalue" , "cowplot" ), dependencies = TRUE)

```

## R Markdown

Welcome to my proteomics benchmarking workflow script! Here you will find a general user's guide to setting up your MS data files and any and all other required files in order to seamlessly run your benchmarking data through this workflow to assess the quantitative accuracy of your own system.

I have included the simple total intensity sum normalization function that we used to normalize our data. We did it in this way in order to only normalize on the spiked-in ecoli peptides or proteins. This was to avoid biasing the data towards the human probes that have an overall higher intensity due to their higher relative concentration compared to the ecolis. Feel free to use this function or normalize externally and comment out the normalization line in the data input chunk.

```{r sum normalization function, echo = FALSE}
sumNorm <- function(diaProt, index) {
tot = 0
# If your dataset has only one species comment out these rows
ecoliRow <- diaProt[which(diaProt$species == "ecoli"), ]
diaProt <- diaProt[which(diaProt$species == "human"), ]
for (i in index:length(colnames(diaProt))) {
  for (j in 1:length(diaProt[,i])) {
    if(is.na(diaProt[j,i])) {
      next
    }
    tot <- tot + diaProt[j,i]
  }
  if (i ==index) {
    group1 <- tot
  }
  fact <- tot/group1
  for (j in 1:length(diaProt[,i])) {
     if(is.na(diaProt[j,i])) {
      next
    }
    diaProt[j,i] <- diaProt[j,i]/fact
  }
  # If your dataset has only one species comment out this for loop
  for(j in 1:length(ecoliRow[,i])) {
    if(is.na(ecoliRow[j,i])) {
      next
    }
    ecoliRow[j,i] <- ecoliRow[j,i]/fact
  }
  tot = 0
}
diaProt <- rbind(diaProt, ecoliRow)
return(diaProt)
}
```

This function is called by main() and performs an fdr correction, assembles confusion matrices, and Reciever-Operating Characteristic curves.

If your dataset contains some species' other than E. coli and human data, there are several string searches below that must be modified to reflect those species (use ctrl-F find/replace)

pi0=1 performs the Benjamini Hochberg correction, for a less stringent option you may set pi0=0 to perform the Storey correction.

```{r Ropeca Confusion Matrix, echo = FALSE}
ropecaConfusion <- function(ttest_data, Norm, Stattest, fdr) {
  
  if (is.null(ttest_data)) {
    return()
  }
  ttest_data$q.value <- NA
  if (Stattest =="t-test") {
    #Q-test
    q <- qvalue(p = ttest_data$p.value, pi0=1)
    ttest_data$q.value <- q$qvalues
  }
  if (Stattest =="Ropeca") {
    #ttest_data$q.value <- ttest_data$p.fdr
    q <- qvalue(p = ttest_data$p, pi0=1)
    ttest_data$q.value <- q$qvalues
  }
  
  if (Stattest =="rots") {
   
     q <- qvalue(p = ttest_data$pvalue, pi0=1)
     ttest_data$q.value <- q$qvalues
  }
  if (Stattest == "Limma") {
     q <- qvalue(p = ttest_data$p.value, pi0=1)
     ttest_data$q.value <- q$qvalues
  }
  if (Stattest=="MSStats") {
    q <- qvalue(p = ttest_data$pvalue, pi0=1)
    ttest_data$q.value <- q$qvalues
  }
  
  if(is.null(ttest_data$species) & Stattest == "rots") {
    ttest_data <- cbind(ttest_data$logfc, q.value = ttest_data$q.value)
    ttest_data <- as.data.frame(ttest_data)
    setDT(ttest_data, keep.rownames=TRUE)[]
    ttest_data[grep("ECOLI", ttest_data$rn), "species"] <- "ecoli"
    ttest_data[grep("HUMAN", ttest_data$rn), "species"] <- "human"
  }
  if(is.null(ttest_data$species) & Stattest == "Ropeca") {
    setDT(ttest_data, keep.rownames = TRUE)[]
    ttest_data[grep("ECOLI", ttest_data$rn), "species"] <- "ecoli"
    ttest_data[grep("HUMAN", ttest_data$rn), "species"] <- "human"
  }
  if(is.null(ttest_data$species) & Stattest == "Limma") {
    ttest_data <- as.data.frame(ttest_data)
    setDT(ttest_data, keep.rownames = TRUE)[]
    ttest_data[grep("ECOLI", ttest_data$rn), "species"] <- "ecoli"
    ttest_data[grep("HUMAN", ttest_data$rn), "species"] <- "human"
    ttest_data[grep("ecoli", ttest_data$rn), "species"] <- "ecoli"
    ttest_data[grep("human", ttest_data$rn), "species"] <- "human"
  }
  
 #Q-value confusion matrix
  confusionQTable <- data.frame("key" = character(0), "value" = numeric(0))

  # True positives DDA
  confusionQTable[1,1] <- "True positives"
  confusionQTable[1,2] <- 
    round(nrow(ttest_data %>% filter(species == "ecoli", q.value < fdr)), 5)
  
  # False negatives DDA
  confusionQTable[2,1] <- "False negatives"
  confusionQTable[2,2] <-
    round(nrow(ttest_data %>% filter(species == "ecoli", q.value >= fdr)), 5)
  
  # False positives DDA
  confusionQTable[3,1] <- "False positives"
  confusionQTable[3,2] <- 
    round(nrow(ttest_data %>% filter(species == "human", q.value < fdr)), 5)
  
  
  # True negatives DDA
  confusionQTable[4,1] <- "True negatives"
  confusionQTable[4,2] <- 
    round(nrow(ttest_data %>% filter(species == "human", q.value >= fdr)), 5)
  
  # E. Coli proteins
  confusionQTable[5,1] <- "Condition positive"
  confusionQTable[5,2] <- round(nrow(ttest_data %>% filter(species == "ecoli")), 5)
  
  # Human proteins
  confusionQTable[6,1] <- "Condition negative"
  confusionQTable[6,2] <- round(nrow(ttest_data %>% filter(species == "human")), 5)
  
  # P < 0.05
  confusionQTable[7,1] <- "Predicted condition positive"
  confusionQTable[7,2] <- round(nrow(ttest_data %>% filter(q.value < fdr)), 5)
  
  # P >= 0.05
  confusionQTable[8,1] <- "Predicted condition negative"
  confusionQTable[8,2] <- round(nrow(ttest_data %>% filter(q.value >= fdr)), 5)
  
  # Total population
  confusionQTable[9,1] <- "Total population"
  confusionQTable[9,2] <- round(nrow(ttest_data), 5)
  
  # Accuracy
  confusionQTable[10,1] <- "Accuracy"
  confusionQTable[10,2] <- round((confusionQTable[1,2] + confusionQTable[4,2]) / confusionQTable[9,2], 4)
  
  # Prevalence
  confusionQTable[11,1] <- "Prevalence"
  confusionQTable[11,2] <- round(confusionQTable[5,2] / confusionQTable[9,2], 4)
  
  # False positive rate
  confusionQTable[12,1] <- "False positive rate"
  confusionQTable[12,2] <- round(confusionQTable[3,2] / confusionQTable[6,2], 4)
  
  # True positive rate
  confusionQTable[13,1] <- "True positive rate"
  confusionQTable[13,2] <- round(confusionQTable[1,2] / confusionQTable[5,2], 4)
  
  # False negative rate
  confusionQTable[14,1] <- "False negative rate"
  confusionQTable[14,2] <- round(confusionQTable[2,2] / confusionQTable[5,2], 4)
  
  # True negative rate
  confusionQTable[15,1] <- "True negative rate"
  confusionQTable[15,2] <- round(confusionQTable[4,2] / confusionQTable[6,2], 4)
  
   # Positive Predicted Value
  confusionQTable[16,1] <- "Positive Predictive Value"
  confusionQTable[16,2] <- round(confusionQTable[1,2] / confusionQTable[7,2], 4)
  
   # Negative Predicted Value
  confusionQTable[17,1] <- "Negative Predictive Value"
  confusionQTable[17,2] <- round(confusionQTable[4,2] / confusionQTable[8,2], 4)
  
  # Print table
  tab <- "q-value Confusion Table"
  plort <- tableGrob(confusionQTable, rows = rownames(confusionQTable), cols =   colnames(confusionQTable), theme = ttheme_default(), vp = NULL)
  grid.arrange(top = paste(Stattest, Norm, tab, sep = " "), plort)

  #Generate ROC curve
  ttest_data$changing <- TRUE
  ttest_data[which(ttest_data$species == "human"), "changing"] <- FALSE
  
  max(ttest_data$q.value)
  x <- seq(0.0, 1.0, 0.01)

  fpr <- c()
  tpr <- c()
  
  cp <- length(which(ttest_data$changing))
  cn <- length(which(!ttest_data$changing))
  for(i in 1:length(x)) {
    ttest_data$test <- FALSE
    ttest_data[which(ttest_data$q.value <= x[i]),"test"] <- TRUE
    
    tpr[i] <- length(which(ttest_data$changing & ttest_data$test)) / cp
    fpr[i] <- length(which(!ttest_data$changing & ttest_data$test)) / cn
  }
  data_roc <- data.frame("threshold" = x, "tpr" = tpr, "fpr" = fpr)
  g  <- ggplot(data_roc, aes(x = fpr, y = tpr, colour = threshold))
  roc <- plot(g + geom_point() + geom_line() +
    theme_pander(base_size = 19) + scale_fill_economist() +
    theme(text=element_text(size=19,family="serif")) +
    theme(axis.ticks = element_blank()) +
    theme(axis.title.y=element_text(margin=margin(0,20,0,0))) +
    theme(axis.title.x=element_text(margin=margin(20,0,0,0))) +
    labs(x="False positive rate", y = "True positive rate") +
    ggtitle(paste(Stattest, Norm, sep= " ")))
  height = (tpr[-1]+tpr[-length(tpr)])/2
  width = diff(fpr)
  AUC <- sum(height*width)
  CM <- slice(confusionQTable, 1,3)
   
  return(list(roc, AUC, CM))
}
```

This is the main section of the analysis script that performs the data analysis, fdr correction, and generates various plots for visualization.


```{r main function, echo = TRUE}
main <- function(input, dataType, quant, cols, fdr, group1, group2, rotGroups, replicates) {
  if (dataType == "protein")   {
    inputLong <- gather(input, run, intensity, 4:cols) %>% filter(!is.na(intensity))
    inputLong$condition <- NA
    inputLong[grep("1x", inputLong$run), "condition"] <- "1x"
    inputLong[grep("2x", inputLong$run), "condition"] <- "2x"
    
    inputLong$quant <- quant
  
    protein_counts <- inputLong %>% group_by(quant, condition, species) %>%
    summarize(count = length(unique(protein)))
  }
  if (dataType == "peptide" & quant == "dia")   {  
    inputLong <- gather(input, run, intensity, 5:cols) %>% filter(!is.na(intensity))
    inputLong$condition <- NA
    inputLong[grep("1x", inputLong$run), "condition"] <- "1x"
    inputLong[grep("2x", inputLong$run), "condition"] <- "2x"
  
    inputLong$quant <- quant
  
    protein_counts <- inputLong %>% group_by(quant, condition, species) %>%
    summarize(count = length(unique(peptide)))
  }
  if (dataType == "peptide" & quant == "dda") {
    inputLong <- gather(input, run, intensity, 4:cols) %>% filter(!is.na(intensity))
    inputLong$condition <- NA
    inputLong[grep("1x", inputLong$run), "condition"] <- "1x"
    inputLong[grep("2x", inputLong$run), "condition"] <- "2x"
  
    inputLong$quant <- quant
    
    protein_counts <- inputLong %>% group_by(quant, condition, species) %>%
    summarize(count = length(unique(peptide)))
  }

      #t-test
       HU_1x <- inputLong %>% filter(species == "human", condition == "1x") %>%
       select(intensity) %>% unlist
       HU_2x <- inputLong %>% filter(species == "human", condition == "2x") %>%
       select(intensity) %>% unlist

       EC_1x <- inputLong %>% filter(species == "ecoli", condition == "1x") %>%
       select(intensity) %>% unlist
       EC_2x <- inputLong %>% filter(species == "ecoli", condition == "2x") %>%
       select(intensity) %>% unlist

       HU <- t.test(x=HU_1x, y=HU_2x)
       EC <- t.test(x=EC_1x, y=EC_2x)

       group_means <- inputLong %>% group_by(species, quant, condition) %>%
       summarize(mean = mean(intensity)) %>% spread(condition, mean)

       names(group_means)[3:4] <- c("1x Mean", "2x Mean")
       group_means[1] <- c("human", "ecoli")

       overall <- inputLong %>% group_by(species, quant) %>% summarize(count = n()) %>% select(-count)

       overall$p.value <- c(HU$p.value, EC$p.value)
       overall <- merge(overall, group_means)
       overall$species <- c("human", "ecoli")
       names(overall)[1:3] <- c("Species", "Quantitation", "P-value")

       overall$Ratio <- overall[, "2x Mean"] / overall[, "1x Mean"]


       overall$lower <- NA
       overall$upper <- NA

       HU_ratio <- ttestratio(HU_2x, HU_1x)
       EC_ratio <- ttestratio(EC_2x, EC_1x)

       overall[2, "lower"] <- EC_ratio$conf.int[[1]]
       overall[2, "upper"] <- EC_ratio$conf.int[[2]]


       overall[1, "lower"] <- HU_ratio$conf.int[[1]]
       overall[1, "upper"] <- HU_ratio$conf.int[[2]]

       overall$expected_ratio <- c(1,2)

      g <- ggplot(overall, aes(x = Quantitation, y = log2(Ratio)))
      plot(g + geom_point(size = 2) + facet_grid(. ~ Species) +
      geom_errorbar(aes(x = Quantitation, ymax = log2(upper), ymin = log2(lower)),
      width = 0.2) +
      geom_point(aes(x = Quantitation, y = log2(expected_ratio)), colour = "blue", size = 3) +
      theme_pander(base_size = 19) + scale_colour_economist() +
      theme(text=element_text(size=19,family="serif")) +
      theme(axis.ticks = element_blank()) +
      theme(axis.title.y=element_text(margin=margin(0,20,0,0))) +
      theme(axis.title.x=element_text(margin=margin(20,0,0,0))) +
      labs(y=expression(Log[2]*Ratio), x = "") +
      scale_y_continuous(breaks=seq(-0.2,1.0,0.2)))

      overall <- overall %>% select(-expected_ratio)
      overall

if (dataType == "protein")   {
  ttest_data <- inputLong %>% group_by(species, quant, protein) %>% do(
    ttest = {
      one <- .[which(.[,"condition"] == "1x"),]
      one <- one$intensity

      two <- .[which(.[,"condition"] == "2x"),]
      two <- two$intensity

      if(length(one) > 1 && length(two) > 1) {
        data_ratio <- ttestratio(two, one)
        
           lower <- data_ratio$conf.int[[1]]
           upper <- data_ratio$conf.int[[2]]
           ratio <- data_ratio$estimate[[3]]
           pratio <- data_ratio$p.value[[1]]

        paste(ratio, pratio, lower, upper, sep = " ")
        }

      else(NA)

      }
    )
}

   if (dataType == "peptide")   {
     ttest_data <- inputLong %>% group_by(species, quant, peptide) %>% do(
       ttest = {
         one <- .[which(.[,"condition"] == "1x"),]
         one <- one$intensity

         two <- .[which(.[,"condition"] == "2x"),]
         two <- two$intensity

         if(length(one) > 1 && length(two) > 1) {
           data_ratio <- ttestratio(two, one)

           lower <- data_ratio$conf.int[[1]]
           upper <- data_ratio$conf.int[[2]]
           ratio <- data_ratio$estimate[[3]]
           pratio <- data_ratio$p.value[[1]]

           paste(ratio, pratio, lower, upper, sep = " ")

           }

         else(NA)

         }
       )
   }
ttest_data <- ttest_data %>% filter(!is.na(ttest)) %>%
  separate(ttest, c("mean", "p.value", "lower", "upper"), " ", convert = TRUE)

ttest_data <- ttest_data %>% filter(!is.na(upper), !is.na(lower), !is.na(p.value))

  elim <-NA
  outLimma <-NA
  namaste <- data.frame()
  if (dataType == "peptide" & quant == "dia")   {
    #  Limma moderated t-test
    limMatrix <- input[-3:-4]
    for (i in 1:length(limMatrix[,1])) {
       namaste[i,1] <- paste(limMatrix[i,1], limMatrix[i,2])
    }
    rownames(limMatrix) <-namaste[,1]
    limMatrix <- limMatrix[,-1:-2]
    designer <- read.delim("lmfit_design_matrix.txt", row.names=1)

    limMatrix <- log(limMatrix, base=2)
    cont.matrix<-makeContrasts(X1x-X2x, levels=designer)
    limmer <- lmFit(limMatrix, designer, method= "ls")
    fit2<-contrasts.fit(limmer, cont.matrix)
    #  Moderated t-statistic
    outLimma <- eBayes(fit2,stdev.coef.lim= c(0.1,1.0))
    
    #elim <- qqt(outLimma$t,df=outLimma$df.prior+outLimma$df.residual,main="Moderated t")
  }
  
  if (dataType == "protein" & quant == "dia")   {
    #  Limma moderated t-test
    limMatrix <- input[-2:-3]
    namaste <- limMatrix[,1]
    rownames(limMatrix) <-namaste
    limMatrix <- limMatrix[-1]
    designer <- read.delim("lmfit_design_matrix.txt", row.names=1)
    
    limMatrix <- log(limMatrix, base=2)
    cont.matrix<-makeContrasts(X1x-X2x, levels=designer)
    limmer <- lmFit(limMatrix, designer, method= "ls")
    fit2<-contrasts.fit(limmer, cont.matrix)
    
    #  Moderated t-statistic
    outLimma <- eBayes(fit2,stdev.coef.lim= c(0.1,1.0))
    #elim <- qqt(outLimma$t,df=outLimma$df.prior+outLimma$df.residual,main="Moderated t")
  }

  if (quant == "dda")   {
    #  Limma moderated t-test
    limMatrix <- input[-2]
    for (i in 1:length(limMatrix[,1])) {
       namaste[i,1] <- paste(limMatrix[i,1], limMatrix[i,2])
    }
    rownames(limMatrix) <-namaste[,1]
    limMatrix <- limMatrix[-1:-2]
    designer <- read.delim("lmfit_design_matrix_dda.txt", row.names=1)
    
    limMatrix <- log(limMatrix, base=2)
    cont.matrix<-makeContrasts(X1x-X2x, levels=designer)
    limmer <- lmFit(limMatrix, designer, method= "ls")
    fit2<-contrasts.fit(limmer, cont.matrix)
    
    #  Moderated t-statistic
    outLimma <- eBayes(fit2,stdev.coef.lim= c(0.1,1.0))
    #elim <- qqt(outLimma$t,df=outLimma$df.prior+outLimma$df.residual,main="Moderated t")
  }
  
 

   peca <- NA
  # ROPECA #
  if((dataType == "peptide") & quant== "dia") {
    pecaDat <- input
    pecaDat$peptide <-paste(pecaDat$peptide,pecaDat$species, sep= "_")
    pecaDat <- pecaDat[,-3]
    pecaDat <- pecaDat[,-3]
    peca <- PECA_df(df = pecaDat, id = "protein", samplenames1= group1, samplenames2= group2, normalize = FALSE, test = "rots", type = "median", progress = TRUE, paired = FALSE)
  }

   rots <- NA
  # # Rots #
  if (dataType == "protein") {
    
    rotD <- input[,-2]
    rotD <- rotD[,-2]
    rotDat <- rotD[,-1]
    rownames(rotDat) <- rotD[,1]
    tot = 0
    i=1
    while(i < length(rownames(rotDat))+1) {
      for (j in 1:length(rotDat[i,])) {
        if(is.na(rotDat[i,j])) {
          tot <- tot + 1
        }
      }
      if (tot > 1) {
       rotDat <- rotDat[-i,]
       i <- i - 1
      }
      tot=0
      i <- i+1
    }
    rots <- ROTS(data = rotDat, groups = rotGroups, B = 1000, K = 4000, log=FALSE)
    summary(rots, fdr = fdr)
  }
   ttes <- ropecaConfusion(ttest_data, replicates, "t-test", fdr)
   tAUC <- ttes[2]
   troc <- ttes[1]

   lm <- ropecaConfusion(outLimma, replicates, "Limma", fdr)
   lmAUC <- lm[[2]]
   lmroc <- lm[1]
   
   roAUC <- NA
   roroc <- NA
   if (dataType == "peptide" & quant == "dia")   {
     ropeca <- ropecaConfusion(peca, replicates, "Ropeca", fdr)
     roAUC <- ropeca[2]
     roroc <- ropeca[1]
   }
   if (dataType == "protein")   {
     ropeca <- ropecaConfusion(rots, replicates, "rots", fdr)
     roAUC <- ropeca[2]
     roroc <- ropeca[1]
   }
   if (dataType=="peptide" & quant == "dia") {
     data_CM <- data.frame("Test"=ttes[[3]]$key, "T-Test" = ttes[[3]]$value, "Limma" =      lm[[3]]$value, "Ropeca" = ropeca[[3]]$value)
   }
   else if (dataType=="protein") {
     data_CM <- data.frame("Test"=ttes[[3]]$key, "T-Test" = ttes[[3]]$value, "Limma" = lm[[3]]$value,  "Rots" = ropeca[[3]]$value)
   }
   else {
     data_CM <- data.frame("Test"=ttes[[3]]$key, "T-Test" = ttes[[3]]$value, "Limma" = lm[[3]]$value)
   }
   
   ex <- data_CM[,1]
   data_CM <- data_CM[,-1]
   rownames(data_CM) <- ex
   data_CM %>% gather()
   CM <- gather(data_CM, Test, Count)
   for(i in 1:length(rownames(CM))) {
     if (i%%2 == 1) {
         CM[i,3] = "True Positives"
       }
     if (i%%2 == 0) {
       CM[i,3] = "False Positives"
     }
   }
  colnames(CM) <- c("Test", "Count", "Label")
  
  return(list(peca,rots,outLimma,ttest_data,tAUC, roAUC, lmAUC, CM))
}

```

Takes DDA input files directly out of MaxQuant and put thems into the correct format to run through the analysis workflow. 

```{r DDA Input, echo = FALSE}

#Load data
dda <- read.csv("Gygi_JPR_2018_LFQ_peptide.csv")
conditions <- read.csv("annotation_dda.csv")
names(conditions) <- c("run", "condition")

#Select your data columns (see sample data if you're unsure of format)
dda <- dda[, c(1, 36, 67:73)]
names(dda)[1] <- "peptide"
names(dda)[2] <- "protein"


dda <- gather(dda, run, intensity, 3:9, convert = TRUE, factor_key = TRUE)
dda$run <- as.character(dda$run)
dda <- merge(dda, conditions)
rm(conditions)


dda$quant <- "dda"
dda$condition <- do.call(paste, c(dda[,c("run", "condition", "quant" )], sep = "_"))

# Remove unneccesary 'Run' column
dda<- dda[,-1]

#Modify to your species of interest
dda$species <- NA
dda[grep("HUMAN", dda$protein), "species"] <- "human"
dda[grep("ECOLI", dda$protein), "species"] <- "ecoli"
dda <- dda[which(!is.na(dda$species)), ]

# Rolls up individual peptide charges into single peptide intensities
ddaPep <- ddply(dda, .(peptide, protein, condition, species, quant), numcolwise(sum,na.rm=TRUE))
# Rolls up peptide species into protein intensities
ddaProt <- ddply(dda, .(protein, condition, species, quant), numcolwise(sum,na.rm=TRUE))

#Replace zeros with NAs
ddaPep[ddaPep==0] <- NA
ddaProt[ddaProt==0] <- NA

#Convert back to wide format
ddaPep <- spread(ddaPep, condition, intensity)
ddaProt <- spread(ddaProt, condition, intensity)

#Normalization via sumNorm function
ddaPep <- sumNorm(ddaPep, 5)
ddaProt <- sumNorm(ddaProt, 4)

# write.csv(ddaPep, "dda_peptide_wide.csv", row.names = F)
# write.csv(ddaProt, "dda_protein_wide.csv", row.names = F)
```

Takes DIA input files directly out of Spectronaut and put thems into the correct format to run through the analysis workflow.

```{r DIA Input, echo = FALSE}
dia <- read.csv("20200114_112622_EA100915_DIA_Benchmark_Report.csv")

# Conditions file should be table of column headers in one column with their corresponding sample group in the adjacent column
conditions <- read.csv("annotation.csv")
names(conditions) <- c("run", "condition")

# Label all non-sample columns
names(dia)[1:3] <- c("peptide", "protein", "Fasta")

dia$quant <- "dia"

# Rearrange all character columns to be if front of MS signals
dia <- dia[,c(1,2,20,4:19)]
dia[dia==0] <- NA


# Run if your data frame is showing numerical columns as containing character objects
diaPep <- mutate_all(dia, function(x) as.numeric(as.character(x)))
diaPep[,1:3] <- dia[,1:3]

# Run if you are doing peptide level quantification with charge state
diaPep[,c(1,20)] <- do.call(rbind, strsplit(as.character(diaPep[,1]),"\\."))

#rRmove charge state
diaPep <- diaPep[,-20]

#Modify to species' present in your dataset
diaPep$species <- NA
diaPep[grep("HUMAN", diaPep$protein), "species"] <- "human"
diaPep[grep("ECOLI", diaPep$protein), "species"] <- "ecoli"
diaPep <- diaPep[which(!is.na(diaPep$species)), ]

# Rearrange
diaPep <- diaPep[,c(1,2,3,20,4:19)]

dia <- gather(diaPep, run, intensity, 5:20, convert = TRUE, factor_key = TRUE)

dia$run <- as.character(dia$run)
dia <- merge(dia, conditions)
rm(conditions)


dia$condition <- do.call(paste, c(dia[,c("run", "condition", "quant" )], sep = "_"))

# Remove unneccesary 'Run' column
dia<- dia[,-1]

# Combine ion fragments into single peptide values and peptides into protein values 
diaPep <- ddply(dia, .(peptide, protein, species, quant, condition), numcolwise(sum,na.rm=TRUE))
diaProt <- ddply(dia, .(protein, species, quant, condition), numcolwise(sum,na.rm=TRUE))

#Convert back to wide format
diaPep <- spread(diaPep, condition, intensity)
diaProt <- spread(diaProt, condition, intensity)

diaPep[diaPep==0] <- NA
diaProt[diaProt==0] <- NA

# Adjust the following integer (index column) to be first column with MS signal
diaPep <- sumNorm(diaPep, 5)

diaProt <- sumNorm(diaProt, 4)

# write.csv(diaPep, "dia_peptide_wide.csv", row.names = F)
# write.csv(diaProt, "dia_protein_wide.csv", row.names = F)
```

This final chunk is for the analysis script setup, sample group layout, and for building several post-analysis plots for visualization.


```{r Matrix Spliting, ref.label = "main function", echo = FALSE}

# Sample groups for ROTS analysis
# Set to NA if Phospho/Peptide Level data
rotGroups <- c(0,0,0,1,1,1,1)


# Sample groups for Ropeca analysis
# Set to NA if not DIA Peptide data

# Example format: group1 = c("EA100915_05_1x_dia", "EA100915_09_1x_dia", "EA100915_13_1x_dia")
#                 group2 = c("EA100915_06_2x_dia", "EA100915_10_2x_dia", "EA100915_14_2x_dia")

group1 <- NA
group2 <- NA


# Choose input file

# input <- read.csv("dia_protein_wide.csv")
# input <- ddaProt


# Write output figures to a pdf
#pdf(file = "DIA_Test.pdf")

# Arguments for main script (1: input matrix, 2: ("peptide" or "protein") quant level, 3: Quant tyoe, 4: Length of input matrix, 5: FDR cutoff, 6,7: Group 1 & 2 for Ropeca analysis, 8: Grouping vector for Rots analysis, 9: Number of Technical Replicates ran)

data <- main(input, "protein", input$quant[1], length(colnames(input)), 0.05, group1, group2, rotGroups, replicates)

# Returns all post-test data matrices, area under curve for the ROC curves, and a slice of the confusion table of true/false positives for each test for making a bar graph

# Generate a volcano plot of 

volcanoDF <- as.data.frame(data[[4]])
volcanoDF$mean <- volcanoDF$mean-1
q <- qvalue(p = volcanoDF$p.value)
volcanoDF$q.value <- q$qvalues
labols <- volcanoDF[,1]
colors <- NULL

i=1
for (i in 1:length(labols)) {
  if (labools[i] == "ecoli") {
    colors[i] = "royalblue"
  }
  if (labools[i] == "human") {
    colors[i] = "red2"
  }
}
custom <- setNames(as.list(colors), labols)
EnhancedVolcano(volcanoDF,
      lab = "",
      x = 'mean',
      y = 'q.value',
      xlim = c(-5, 8),
      ylim = c(0,10),
      FCcutoff = 1.0,
      pCutoff=0.05,
      title= "DIA Protein MS1",
      subtitle = "8 Replicates",
      colCustom = custom)



# Generate a bar graph of true/false positives for each statistical test

g <- ggplot(data[[8]], aes(x=Test, y=Count, fill = Label))
  plot(g +geom_bar(stat="identity", position="stack" ) +
          theme_pander(base_size = 15) + scale_fill_economist() +
          scale_y_continuous(limits = c(0,900)) +
          ggtitle("DDA Protein Gygi"))

  
  

dev.off
```




```{r , echo = FALSE}

```

